Mahotas: Open source software for scriptable computer vision

### 2012-08-29 paper.tex
TXT Machine Learning Packages

The interface is a procedural interface, with most functions working
independently of each other (there is code sharing at the implementation
level, but this is hidden from the user).

@@

\begin{description}
\item[\textsc{surf}] Speeded-up Robust Features~\citep{eth_biwi_00517}. This is
a collection of local features.
\item[features] Global feature descriptors. In particular, Haralick texture
features, Zernike moments, local binary patterns, threshold adjacency
statistics.
\item[Wavelet] Haar and Daubechies wavelets. Forward and inverse transforms.
\item[morphological functions] Erosion and dilation, as well as some more
complex operations built on these. There are both binary and grayscale
implementations of these operators.
\item[watershed] seeded watershed and distance map
transforms~\citep{felzenszwalb}.
\item[filtering] Gaussian filtering and general convolutions.
\item[polygon operations] convex hull, polygon drawing.
\end{description}

In general, functions operate on all data types (except when it not well
defined). This is performed without any extra memory copies. Mahotas is heavily
optimised for both speed and memory usage.

@@

Often, functions take an argument named \code{out} where the output will be
stored. This argument is often much more restricted in type. In particular, it
must often be a contiguous array.\footnote{Numpy supports non-contiguous
arrays, which are most often slices into other, larger, contiguous arrays
(e.g., given a $128 \times 128$ contiguous array, one can build a $64 \times
128$ non-contiguous array by taking every other row).} Since this is a
performance feature, it is natural that the interface is less flexible
(accessing a contiguous array is much more efficient than a non-contiguous
one).

@@

This is a simple example of mahotas usage. Code for this and other examples is
present in the mahotas source distribution under the \texttt{demos/} directory.

In this tutorial, we load an image and find SURF interest point and descriptors
on it.

@@

The first step is to load the image and convert to 8~bit numbers.
\begin{python}
f = mahotas.imread('luispedro.jpg', as_grey=True)
f = f.astype(np.uint8)
\end{python}

@@

\begin{python}
descrs = spoints[:,6:]

values, _ = milk.kmeans(descrs, 5)
\end{python}

Finally, we can show the 
\begin{python}
colors = np.array(
    [ 255,  25,   1],
    [203,  77,  37],
    [151, 129,  56],
    [ 99, 181,  52],
    [ 47, 233,   5]])
f2 = surf.show_surf(f, spoints[:64], values, colors)
\end{python}

Using matplotlib~\citep{10.1109/MCSE.2007.55}, we finally display the image as
Figure~\ref{fig:surf}.

@@

Mahotas is written in \cpp, but this is completely hidden from the user as
there are hand-written Python wrappers for all functions (automatically
generated wrappers inevitably lead to worse error messages and are less
flexible).

@@

Mahotas uses the standard Python build machinery and distribution channels.
Building and installing from source code is done using
\begin{verbatim}
python setup.py install
\end{verbatim}
Alternatively, Python based package managers (such as \texttt{easy\_install} or
\texttt{pip}) can be used (mahotas works well with these systems).

@@

\textbf{List of contributors}\\
Luis Pedro Coelho (Carnegie Mellon University and Instituto de Medicina
Molecular), Zachary Pincus (Stanford University), Peter J. Verveer (European
Molecular Biology Laboratory), Davis King (Northrop Grumman ES), Robert Webb
(Carnegie Mellon University), Matthew Goodman (University of Texas at Austin),
K.-Michael Aye (University of Bern), Rita Sim\~{o}es (University of Twente),
Joe Kington (University of Wisconsin Dept.  of Geoscience), Christoph Gohlke
(University of California, Irvine), and Sandro Knauss (University of Bremen).

@@

Mahotas does not include machine learning related functionality, such as
$k$-means clustering or any classification. This is the result of an explicit
design decision. Specialised machine learning packages for Python (e.g.,
milk\footnote{http://github.com/luispedro/milk},
scikit-learn\footnote{http://scikit-learn}) already
exist~\citep{Pedregosa:2011:SML:2078183.2078195}. A good classification (or
other) system can benefit both computer vision users and others. As these
projects all use Numpy arrays as their data types, it is easy to use
functionality from the different project seamlessly (no copying of data is
necessary).

>>>>

The interface is a procedural interface, with no global state. All functions
work independently of each other (there is code sharing at the implementation
level, but this is hidden from the user).

@@

\begin{description}
\item[\textsc{surf}] Speeded-up Robust Features~\citep{eth_biwi_00517}. This
includes both keypoint detection and descriptor computation.
\item[features] Global feature descriptors. In particular, Haralick texture
features, Zernike moments, local binary patterns, and threshold adjacency
statistics (both the original~\citep{Hamilton2007} and the parameter-free
versions~\citep{Coelho2010}).
\item[Wavelet] Haar and Daubechies wavelets. Forward and inverse transforms.
\item[morphological functions] Erosion and dilation, as well as some more
complex operations built on these. There are both binary and grayscale
implementations of these operators.
\item[watershed] seeded watershed and distance map
transforms~\citep{felzenszwalb}.
\item[filtering] Gaussian filtering and general convolutions.
\item[polygon operations] convex hull, polygon drawing.
\end{description}

In general, functions operate on all data types (in some cases, the operation
is only meaningful for floating point arrays, for example). This is performed
without any extra memory copies. Mahotas is heavily optimised for both speed
and memory usage (it can be used with very large arrays).

@@

When a new image is to be returned, functions take an argument named \code{out}
where the output will be stored. This argument is often much more restricted in
type. In particular, it must often be a contiguous array.\footnote{Numpy
supports non-contiguous arrays, which are most often slices into other, larger,
contiguous arrays (e.g., given a $128 \times 128$ contiguous array, one can
build a $64 \times 128$ non-contiguous array by taking every other row).} Since
this is a performance feature (its purpose is to avoid extra memory
allocation), it is natural that the interface is less flexible (accessing a
contiguous array is much more efficient than a non-contiguous one).

@@

This is a simple example of mahotas usage. Code for this and other examples is
present in the mahotas source distribution under the \texttt{demos/} directory.
In this example, we load an image and find SURF interest point and descriptors
on it.

@@

The first step is to load the image and convert to 8~bit numbers. In this
case, the conversion is done using standard numpy methods, namely
\code{astype}:

\begin{python}
f = mahotas.imread('luispedro.jpg', as_grey=True)
f = f.astype(np.uint8)
\end{python}

@@

\begin{python}
descrs = spoints[:,6:]
values, _ = milk.kmeans(descrs, 5)
\end{python}

Finally, we can show the points in colour.
\begin{python}
colors = np.array(
    [ 255,  25,   1],
    [203,  77,  37],
    [151, 129,  56],
    [ 99, 181,  52],
    [ 47, 233,   5]])
f2 = surf.show_surf(f, spoints[:64], values, colors)
\end{python}

The \texttt{show\_surf} only builds the image as a multi-channel (one for each
colour) image. Using matplotlib~\citep{10.1109/MCSE.2007.55}, we finally
display the image as Figure~\ref{fig:surf}.

@@

Mahotas is mostly written in \cpp, but this is completely hidden from the user
as there are hand-written Python wrappers for all functions (automatically
generated wrappers inevitably lead to worse error messages and are less
flexible).

@@

In keeping with the philosophy of blending in with the ecosystem, Mahotas uses
the standard Python build machinery and distribution channels. Building and
installing from source code is done using
\begin{verbatim}
python setup.py install
\end{verbatim}
Alternatively, Python based package managers (such as \texttt{easy\_install} or
\texttt{pip}) can be used (mahotas works well with these systems).

@@

\textbf{List of contributors}\\
Luis Pedro Coelho (Carnegie Mellon University and Instituto de Medicina
Molecular), Zachary Pincus (Stanford University), Peter J. Verveer (European
Molecular Biology Laboratory), Davis King (Northrop Grumman ES), Robert Webb
(Carnegie Mellon University), Matthew Goodman (University of Texas at Austin),
K.-Michael Aye (University of Bern), Rita Sim\~{o}es (University of Twente),
Joe Kington (University of Wisconsin), Christoph Gohlke (University of
California, Irvine), and Sandro Knauss (University of Bremen).

@@

Mahotas does not include machine learning related functionality, such as
$k$-means clustering or any classification. This is the result of an explicit
design decision. Specialised machine learning packages for Python already
exist~\citep{Pedregosa:2011:SML:2078183.2078195,springerlink:10.1007/978-3-540-30116-5_58,Schaul:2010:PYB:1756006.1756030,Sonnenburg:2010:SML:1756006.1859911}.
A good classification (or other) system can benefit both computer vision users
and others. As these projects all use Numpy arrays as their data types, it is
easy to use functionality from the different project seamlessly (no copying of
data is necessary).
